---
title: "Lab 1"
author: "Srinidhi H"
date: "Week 1 Lab"
output: html_document
---

### 1. Changing the author field and file name.

##### (a) Change the `author:` field on the Rmd document from Your Name Here to your own name.

##### (b) Rename this file to "lab01_YourHameHere.Rmd", where YourNameHere is changed to your own name.

### 2.  Hello World!

Here's an R code chunk that prints the text 'Hello world!'.

```{r}
print("Hello world!")
```

#### (a) Modify the code chunk below to print your name
```{r}
# Edit me

```

<br>

### 3.  Creating sequences
We just learned about the `c()` operator, which forms a vector from its arguments.  If we're trying to build a vector containing a sequence of numbers, there are several useful functions at our disposal.  These are the colon operator `:` and the sequence function `seq()`.

##### `:` Colon operator: 
```{r}
1:10 # Numbers 1 to 10
127:132 # Numbers 127 to 132
```

##### `seq` function: `seq(from, to, by)`
```{r}
seq(1,10,1) # Numbers 1 to 10
seq(1,10,2) # Odd numbers from 1 to 10
seq(2,10,2) # Even numbers from 2 to 10
```

> To learn more about a function, type `?functionname` into your console.  E.g., `?seq` pulls up a Help file with the R documentation for the `seq` function.

#### (a) Use `:` to output the sequence of numbers from 3 to 12
```{r}
# Edit me
3:12
```

#### (b) Use `seq()` to output the sequence of numbers from 3 to 30 in increments of 3
```{r}
# Edit me
seq(3,30,3)
```

#### (c) Save the sequence from (a) as a variable `x`, and the sequence from (b) as a variable `y`.  Output their product `x*y`
```{r}
# Edit me
x = 3:12
y = seq(3,30,3)
print(x*y)
```

<br>

### 4. Cars data

We'll look at data frame and plotting in much more detail in later classes.  For a previous of what's to come, here's a very basic example.

For this example we'll use a very simple dataset.  The `cars` data comes with the default installation of R.  To see the first few columns of the data, just type `head(cars)`.

```{r}
head(cars)
```

We'll do a bad thing here and use the `attach()` command, which will allow us to access the `speed` and `dist` columns of `cars` as though they were vectors in our workspace.

```{r}
attach(cars) # Using this command is poor style.  We will avoid it in the future.
speed
dist
```

#### (a) Calculate the average and standard deviation of speed and distance.
```{r}
# Edit me
mean(speed)
mean(dist)
sd(speed)
sd(dist)
```

<br>

<br>  We can easily produce a histogram of stopping distance using the `hist` function.
```{r}
hist(dist) # Histogram of stopping distance
```

The `plot(x,y,...)` function plots a vector `y` against a vector `x`.  You can type `?plot` into the Console to learn more about the basic plot function. 

#### (b) Use the `plot(x,y)` function to create a scatterplot of dist against speed.
```{r}
# Edit me
plot(speed,dist,type="l")
```

### 5. Problem 1
Assume that we have registered the height and weight for four people: Heights in cm are 180, 165, 160, 193; weights in kg are 87, 58, 65, 100. Make two vectors, height and weight, with the data. 

		The bodymass index (BMI) is defined as
				weight in kg/ (height in m)2

	Make a vector with the BMI values for the four people, and a vector with the 	natural logarithm to the BMI values. Finally make a vector with the weights for 	those people who have a BMI larger than 25.
	
```{r}
#Edit me
height <- c(180, 165, 160, 193) / 100
weight <- c(87, 58, 65, 100)
bmi <- weight / height ^ 2
bmi
bmi25 <- bmi[bmi > 25]
bmi25
log(bmi25)
```

### 6. Problem 2

Assume that we have the following three observations of temperature in celsius: 23, 27, 19. Make a vector with these values. Recall the relation between the Celcius and Fahrenheit temperature scale:
	degress in Fahrenheit = degrees in Celcisus *9/5 +32

Make a new vector with the temperatures in Fahrenheit.

```{r}
#Edit me
temp <- c(23, 27, 19)
fah <- temp * 9 / 5 + 32
fah
```

### 7. Problem 3

Assume that you are interested in cone-shaped structures, and have measured the height and radius of 6 cones. 

Make vectors with these values as follows:
R <- c(2.27, 1.98, 1.69, 1.88, 1.64, 2.14)
H <- c(8.28, 8.04, 9.06, 8.70, 7.58, 8.34)

Recall that the volume of a cone with radius R and height H is given by 1/3*pi*R^2*H. Make a vector with the volumes of the 8 cones.

	Compute the mean, median and standard deviation of the cone volumes. 	Compute also the mean of volume for the cones with a height less than 8.5.
```{r}
R <- c(2.27, 1.98, 1.69, 1.88, 1.64, 2.14)
H <- c(8.28, 8.04, 9.06, 8.70, 7.58, 8.34)
vol <- pi/3 * R ^2 * H
vol
mean(vol)
median(vol)
sd(vol)
```
